Novelty Detection Using Graphical Models for Semantic Room Classification

### 2011-05-16 semantic-nd/article.tex
Changes on Related Work section

Its with a final interest on integrated the developed methods on Dora that this
initial work focus on the novelty detection using probabilistic graphical models
that model the conceptual knowledge and information handled by Dora.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\cite{quattoni2009recognizing} showed that most scene recognition models work poorly in indoor
scenes when compared to outdoor scenes results.
Since the properties that characterize rooms changes conforming its category. Namely corridors are
well described by global properties and bookstores are well described by the presence of specific objects (books).
Their work shows a multi-modal approach is expected to yield better performance by mixing several sources of information.

@@

Probabilistic representations are used in several localised functions in robots operating in the real-world~\cite{gross2009toomas,maierprobabilistic}. And some employ, up to some extent, a probabilistic representation across some subsystems~\cite{kraft2008exploration}.
\cite{vasudevan2008bayesian} performed room categorization through Bayesian reasoning about the presence of objects but did not included observations models (perception is considered deterministic).
And \cite{boutell2006factor} have studied outdoor scene classification using \emph{factor graphs} and modelling spatial relations between objects in the scene to extract better knowledge from semantic (high-level) features.

Its expected that using a unified probabilistic model from the whole system, such as \cite{pronobis2011exploiting}, more information can be reused to correctly predict a given random variable.

Although this paper deals with novelty detection on the robotics area, it does so using very
standard concepts and techniques such as semantic data and graphical models.
Those are often found on areas related with information retrieval.
Interesting examples are works in automatic image annotation using an hidden concept layer between visual features and text information\cite{zhang2005probabilistic}.

@@

A \emph{factor graph} is a bipartite graph connecting a set of nodes representing
random variables and factors $\phi_G$.
Each factor represents a function dependent only on the variables to where its connected.
As so given factor graph can be seen as description of a distribution function over a set
of variables obtained by the product of all the factors. In order to represent a probability
a normalization factor needs to be introduced:

\begin{equation}
P_G(x) = \frac{1}{Z}\prod_{}{\phi(...)},\qquad Z = \sum\prod{\phi(...)}
\end{equation}

Describing the distribution function in terms of graphs allows to use algorithms
such as the Sum-Product~\citep{kschischang2001factor} to efficiently calculate
marginals on any given subset $x$ of variables by exploiting conditional
independence between variables.

\begin{equation}
P_G(x) = \frac{1}{Z}\sum_{V \\ x}{\prod{\phi(...)}}
\end{equation}

>>>>

Its with a final interest on integrating the developed methods on Dora that this
initial work focus on the novelty detection using probabilistic graphical models
that model the conceptual knowledge and information handled by Dora.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\cite{quattoni2009recognizing} showed that most scene recognition models work poorly
in indoor scenes when compared to outdoor scenes results since the properties that
characterize rooms changes conforming its category. Namely corridors are well
described by global properties and bookstores are well described by the presence of
specific objects (books). Their work shows a multi-modal approach is expected to
yield better performance by mixing several sources of information.

@@

Probabilistic representations are used in several localised functions in robots
operating in the real-world~\cite{gross2009toomas,maierprobabilistic}.
And some employ, up to some extent, a probabilistic representation across some
subsystems~\cite{kraft2008exploration}.
And \cite{boutell2006factor} have studied outdoor scene classification using
\emph{factor graphs} and modelling spatial relations between objects in the scene
to extract better knowledge from semantic (high-level) features.

% Its expected that using a unified probabilistic model from the whole system,
% such as \cite{pronobis2011exploiting}, more information can be reused to correctly
% predict a given random variable.

Although this paper deals with novelty detection on the robotics area, it does so
using standard concepts and techniques such as semantic data and graphical models.
Those are often found on areas related with information retrieval.
An interesting example is the work in automatic image annotation using an hidden
concept layer between visual features and text information by \cite{zhang2005probabilistic}.

@@

A \emph{factor graph} is a bipartite graph connecting a set of nodes representing
random variables and factors.
Each factor represents a function dependent only on the variables to where its connected.
As so given factor graph can be seen as description of a distribution function over a set
of variables obtained by the product of all the factors. In order to represent a probability
a normalization factor needs to be introduced:

\begin{equation}
P_G(x) = \frac{1}{Z}\prod_{\phi \in F_G}{\phi(x_{\phi})},\qquad
Z = \sum_{x \in X_G}\prod_{\phi \in F_G}{\phi(x_{\phi})}
\end{equation}

Describing the distribution function in terms of graphs allows to use algorithms
such as the Sum-Product~\citep{kschischang2001factor} to efficiently calculate
marginals on any given subset $x_m$ of variables by exploiting conditional
independence between variables.
